---
title: "Get started with netaipq"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with netaipq}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## What netaipq is for

`netaipq` is the [pqverse](https://github.com/adrientaudiere) home for
**networks and machine-learning** analyses of
[`phyloseq`](https://joey711.github.io/phyloseq/) objects: co-occurrence and
association networks, bipartite-network visualisation, supervised
classification, indicator-species and differential-abundance methods, and
other *genuinely new analysis methods* that go beyond a simple `ggplot2`
wrapper.

By absorbing any feature that would add a brand-new machine-learning, network,
graph, or statistical-analysis dependency, `netaipq` keeps
[`MiscMetabar`](https://adrientaudiere.github.io/MiscMetabar/) lean. Pure
`ggplot2` visualisation belongs in `ggplotpq`; cross-study comparators belong
in `comparpq`.

## Installation

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("adrientaudiere/netaipq")
```

## A worked example: bipartite networks

A bipartite network links two kinds of nodes — here **samples** and **taxa** —
with an edge whenever a taxon occurs in a sample. `bipartite_network_pq()`
builds and draws this network from a `phyloseq` object.

For a readable figure we take a few samples and the commonest taxa from
`data_fungi_mini` (shipped with `MiscMetabar`).

```{r setup, message = FALSE}
library(netaipq)
data(data_fungi_mini)

ps <- prune_samples(sample_names(data_fungi_mini)[1:5], data_fungi_mini)
ps <- clean_pq(ps, silent = TRUE)
top_taxa <- names(sort(taxa_sums(ps), decreasing = TRUE))[1:20]
ps <- prune_taxa(top_taxa, ps)
ps
```

```{r network}
bipartite_network_pq(ps, taxa_color = "Phylum", seed = 1)
```

Taxa nodes are coloured by their `Phylum`; edges connect each taxon to the
samples it was observed in. The function returns a `ggplot` object, so you can
post-process it with any `ggplot2` layer or theme — for example those provided
by `ggplotpq`.

## Where to go next

- The [function reference](https://adrientaudiere.github.io/netaipq/reference/index.html)
  lists every argument of `bipartite_network_pq()`, including node sizing,
  layout, and label control.
- `netaipq` is in active development: network, machine-learning, and
  causal-inference methods are being added. See the
  [ROADMAP](https://github.com/adrientaudiere/pqverse/blob/master/ROADMAP.md)
  for what is coming next.
